### 2025-04-13 fabryka/report_3.Rmd
project and git control version creation

---
title: "Fabryka report"
date: "`r Sys.Date()`"
output: pdf_document
editor_options:
  markdown:
    wrap: 72
params:
  n: NA
---

```{r setup, echo = FALSE, message = FALSE, include = FALSE, warning = FALSE}

suppressPackageStartupMessages({
  library(ggplot2)
})

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

This report is made thanks to Fabryka 1.0.

# Tutorial
## Uploading your data

First, upload your .csv file in the left sidebar panel. Pay attention to your decimal and column delimiters.

Fabryka supports three angle measurement methods:

- angles taken with a compass and an inclinometer
- angles taken with a DistoX2
- angles taken with two shots at the total station.

Five cases (i.e. forms of data) are covered by the application:

- check 'Case 1: Only angles' if you have in your dataset the columns 'id', 'sample', 'code', direction' and 'dip'.
- check 'Case 2: Only angles from DistoX2' if you have in your dataset the same columns that above but your data comes from a DistoX2 and needs to be converted.
- check 'Case 3: Angles from DistoX2 with coordinates' if you have in your dataset the same columns that above (Case 2, data from DistoX2) with a single set of coordinates (x, y, z).
- check 'Case 4: Angles and coordinates' if your data has the same form that above (Case 3) but is coming from a compass and a clinometer.
- check 'Case 5: Two shots data without angles' if you have the columns 'id', 'sample', 'code' and the angles have been measured with a total station (i.e. two shots data).

If you miss a context column, please create it by repeating a same character.

According to your method of angle measurement and the presence of coordinates you have to select the right case. Please refer to the example files in the left side bar panel. It contains a dataset coming from Le Moustier Lower shelter (Dordogne, France). The exmample files contain directly the good column names. You can use them directly in the application. Whatever the names of the columns in your file, you can select the appropriate columns in the column selector. Please keep only complete lines (i.e. no empty modalities, so no line without a fabric measurement). When a data table with new columns appears below the column selector, you are ready to use the application.

If your data comes from a DistoX2, the columns 'direction' (= 'bearing', 0 to 360° angles) and 'dip' (= 'plunge') are automatically corrected in the new table. Whatever the form of your dataset, two new columns 'orientation_pi' (0 to 180° angles) and 'angle_double' (see Krumbein 1939) appear. If the angles were measured using a total station (i.e. two shots data), the bearing and the plunge are also calculated. If only one set of coordinates is provided with angles, a second set of coordinates is calculated and displayed in the new table. You can download the new data by pushing the 'copy', the 'csv' or the 'pdf' buttons.

## Classical method

The classical method is available in all cases (i.e. source of data and absence or presence of coordinates). You can control the data you plot and the data appearing in the figures and the summary data table by changing the checkboxes filters (samples and code) and the slider (i.e. minimum number of measures) in the left sidebar panel. Certain geological processes can have a different impact on objects depending on their physical properties (e.g. silex versus bones in a water flow as in the example files provided where bones have a strong preferential orientation and are much more 'linear' in the Benn diagram, see Thomas et al., 2019, Texier et al., 2020). The natural processes models that you can add to the Benn diagram are calculated using data from Bertran & Lenoble (2002). In the rose diagram, orientations are considered (and not directions) as the Rayleigh test is perform on orientations (see Curray 1956) in the summary table. You can download all the figures by selecting a file format and pressing the 'download' button.

In this last table:

-  'n' stands for number of measures in the samples
- 'E1', 'E2' and 'E3' are the eigenvalues (see Woodcock 1977). They correspond to the sum of the cosines formed by the axis of the objects and the three axis of the space. They are used to compute the Benn indices.
- 'IS', 'PL' and 'EL' are calculated thanks to the eigenvalues (Benn 1994). These are the Benn indices, ratios standing respectively for, 'isotropy' with IS = E3/E1, 'elongation' with EL = 1-(E2/E1) and 'planar' PL = 1-IS-EL
- 'L' is the vector magnitude (strength of the preferred artefact orientation) and 'R.p' is the p-value result of the Rayleigh test.
- 'L.double' and 'R.p.double' are the same statistics that above for double angles (see Krumbein 1939).

Curray's formula to calculate 'L' is (1.1)

(1.1)$$L =\frac{r*100}{n} $$

with (1.2)

(1.2)$$ r = \sqrt{\sum_{i=\alpha}^n sin{(2\alpha)} + \sum_{i=\alpha}^n cos{(2\alpha)}}  $$

where 'n' is the number of measures and 'α', is the orientation (0 to 180°)

The Rayleigh test is perform thanks to the formula (1.3)

(2)$$p = e^{(-(n*L^2))*10^-4}$$

## Spatialised method

In order to distinguish different statistical populations, a spatial analysis method proposed by McPherron (2018) is used. Thanks to this methods it is possible to calculate the Benn indices and to test the orientation rate L (formula (1.1) modified R McPherron 2018 script computed for orientations and double angles) for each remains, taking into account its n nearest neighbors in 3 dimensions (R McPherron 2018 script, modified , formula (2.)). The analysis of the fabric as proposed by McPherron (2018) has the advantage of spatializing the measurements and thus avoiding the arbitrary sampling classically carried out by square or by décapage. It is possible to localize the statistical signal of a series by varying the number of measurements sampled to calculate the Benn indices. The choice of about 40 to 50 measurements (each vestige and its 40 to 50 closest neighbors) to calculate the Benn indices allows local particularities to be observed, and above all, it allows the Rayleigh test to be carried out in its field of application (stability of the test between 40 and 50 measurements according to Bertran and Lenoble 2002). Please note that the Rayleigh test is sample size sensible (fig. 1). 
Several modes of representation are used in the 'Benn diagram & projections' subpanel.

The Benn indices (1994) calculated for each series of remains are projected within a ternary diagram in which each pole (isotropic, linear and planar) is colored respectively in red, green and blue. The RGB color code is reused to materialize the fabric of the remains on spatial projections of the material along the axes of the excavation grid. Projections (xy, xz and yz axes) are performed. On the latter, the orientation of the objects can be materialized by sticks (by uncecking the corresponding box in the left sidebar panel) which color is representative of the fabric of the series formed by the vestige and its 50 closest neighbors. In the left sidebar panel, you can control the samples taken into account in the analyses by selecting the sample explored or by distinguishing the code used (e.g. flint versus bone). You can also use the slider to change the number of neighbours searched for around each point. Be careful, this will alter the results obtained in the Benn diagram and Rayleigh tests.

At the spatial series scale (i.e. each points and n neighbors), statistics are provided in a table (Spatial series data table subpanel) with the Benn indices, the orientation rates L, L.double and p-values returned by the Rayleigh tests. At the sample scale, mean and standard deviations values of Benn indices, mean L and its corresponding Rayleigh tests means (mean values of all spatial series in the sample) and rejections ratios 'Tr' (formula 2.) are given in the subpanel 'Summary table of samples'.

The rejection ratio 'Tr' is perform thanks to the formula (2.)

(2) $$Tr = \frac{(n times p < 0,05)* 100}{n}$$

## Spatial exploration

In this panel, you can perform a spatialised analysis of your data starting from a Benn diagram or from projections. In both cases, you need to select the data you want to explor more locally in a new analysis. From the projections, you can select the points you want to explore directly. In the Benn diagram, you need to select the lasso selector. Pay attention to the number of points selected and check that the new sample contains enough points to carry out the new analysis. If this is not the case, you can lower the slider in the left sidebar panel to perform the analysis with fewer neighbours. You can dowload the new samples you have selected from the corresponding subpanels.

## Markdown report

This panel is in progress. It will be available in the next version of the application. By pushing the download button you will find only a beta test with the tutorial.

```{r echo = FALSE, message = FALSE, fig.cap="Evolution of the p-value returned by the Rayleigh test as a function of the orientation rate L and the number of measures n."}

require(ggplot2)
L = c(0:60)

rayleigh <- function(L,n){
  p = exp(-(n*L^2)*10^-4)
  p_value <- p
  p_value
}

p_value_sim_30 <- rep(x = NA, times = length (L))
for (i in 1:61) {
  p_value_sim_30[i]<- rayleigh(i,30)
}

p_value_sim_40 <- rep(x = NA, times = length (L))
for (i in 1:61) {
  p_value_sim_40[i]<- rayleigh(i,40)
}

p_value_sim_50 <- rep(x = NA, times = length (L))
for (i in 1:61) {
  p_value_sim_50[i]<- rayleigh(i,50)
}

p_value_sim_100 <- rep(x = NA, times = length (L))
for (i in 1:61) {
  p_value_sim_100[i]<- rayleigh(i,100)
}

seuil_0.05 = c(rep(0.05, length(L)))

rayleigh_table <- data.frame(L, p_value_sim_30, p_value_sim_40, p_value_sim_50, p_value_sim_100,seuil_0.05)

ggplot(rayleigh_table, aes(x=L)) + 
    geom_line(aes(y = p_value_sim_100, colour = "n = 100"), size = 1) +
    geom_line(aes(y = p_value_sim_30, colour = "n =  30"), size= 1) +
    geom_line(aes(y = p_value_sim_40, colour = "n =  40"), size= 1) + 
    geom_line(aes(y = p_value_sim_50, colour = "n =  50"), size= 1) +
    geom_line(aes(y = seuil_0.05, colour = "seuil de 0.05"), size= 0.3) +
    xlab("Taux d'orientation L (%)") +
    ylab("Valeur p du test de Rayleigh") + 
    scale_color_manual(values=c("#0073C2", "#EFC000", "#868686", "#3B3B3B", "red")) +         scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60)) +
    scale_y_continuous(breaks = c(0, 0.10, 0.20, 0.30, 0.40, 0.50, 0.60, 0.70, 0.80, 0.90, 1)) +
  theme_minimal()

```