### 2025-07-14 DEPENDENCIES.Rmd
correction of fabryka article and shiny application after first round of review in PCI archaeology

---
title: "fabryka"
Version: 1.1
output: github_document
---

```{=html}
<style>
body {
text-align: justify}
</style>
```

# [<img src="https://raw.githubusercontent.com/marchaeologist/fabryka/main/inst/www/hex_Fabryka4.png" height="250" align="center"/>](https://github.com/marchaeologist/fabryka)

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.15236695.svg)](https://doi.org/10.5281/zenodo.15236695)

# Package dependencies

Find below a list of all packages used in fabryka.

```{r library_fabryka}

fabryka_packages=c("shiny", "dplyr","ggplot2","plotly", "ggsci", "ggtern", "circular", "CircStats", "Ternary", 
                   "RStoolbox", "shinyjs", "rmarkdown", "shinythemes",  "ggalt", "DT", "raster")
print(fabryka_packages)

```

Find in the code chunk below a way to install all the packages used in fabryka.

```{r install_library, eval = FALSE, tidy=FALSE, width=120}

fabryka_packages=c("shiny", "dplyr","ggplot2","plotly", "ggsci", "ggtern", "circular", "CircStats", "Ternary", 
                   "RStoolbox", "shinyjs", "rmarkdown", "shinythemes",  "ggalt", "DT", "raster")
is_installed=sapply(fabryka_packages,require,character.only=T)
sapply(fabryka_packages[!is_installed],install.packages)

```

Find below a list of all packages dependencies used in fabryka.

```{r dependencies}
options(repos = c(CRAN = "https://cloud.r-project.org"))
dependencies <- tools::package_dependencies(fabryka_packages, db = available.packages())
print(dependencies)
```

Please note that there are currently conflicts between the functions of several packages. You will find a list below. This problem will be addressed more carefully in future versions of the application. For now, all functions have been prefixed by their namespace.

```{r conflicts}
library(conflicted)
conflicted::conflict_scout(fabryka_packages)
```

Below find the list of imported packaged by using the usethis::use_package() command.

```{r imports, eval=FALSE}

usethis::use_package("shiny")
usethis::use_package("dplyr")
usethis::use_package("ggplot2")
usethis::use_package("plotly")
usethis::use_package("ggsci")
usethis::use_package("ggtern")
usethis::use_package("circular")
usethis::use_package("CircStats")
usethis::use_package("Ternary")
usethis::use_package("RStoolbox")
usethis::use_package("shinyjs")
usethis::use_package("rmarkdown")
usethis::use_package("shinythemes")
usethis::use_package("ggalt")
usethis::use_package("DT")
usethis::use_package("raster")

```