### 2025-07-14 README.Rmd

---
title: "fabryka"
Version: 1.1
Author: Marc THOMAS [aut, cre] 
output:
  github_document
---

```{=html}
<style>
body {
text-align: justify}
</style>
```

# [<img src="https://raw.githubusercontent.com/marchaeologist/fabryka/main/inst/www/hex_Fabryka4.png" height="250" align="center"/>](https://github.com/marchaeologist/fabryka)

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.15236695.svg)](https://doi.org/10.5281/zenodo.15236695)

## README

fabryka is an open access R *shiny* user-friendly web application dedicated to the fabric analysis of archaeological or geological assemblage. fabryka uses classical analytical procedures (Benn, rose, Schmidt and Woodcock diagrams and statistical tests on orientations) for analysing the fabrics of objects, an spatial method and new ways of spatial and statistical exploration. A descriptive article is available at <https://zenodo.org/records/15236695>.

## USE fabryka

The web application is available online at <https://marchaeologist.shinyapps.io/fabryka/>.

You can also use fabryka in RStudio. To do so :

* Install [R](https://www.r-project.org) and [Rstudio Desktop](https://posit.co/download/rstudio-desktop/).

* clone the fabryka repository from GitHub: <https://github.com/marchaeologist/fabryka> or download the code in "Code" and "Download ZIP" at the same address. This will give you the latest version of the source code. The application can be launched by opening the app.R file in the main folder and by running the following commands:

```{r, eval=FALSE}

pkgload::load_all(".")
fabryka()

```

This can be useful if you want to use fabryka in the field without an internet access.

## CONTACT

If you have problems, questions or suggestions, please contact me at marcthomas1@hotmail.fr or via the fabryka repository issues page at <https://github.com/marchaeologist/fabryka/issues>.
All contributors to the project are listed in the 'Contributors' panel in the application.

## SOURCE CODE

The source code is available on github: <https://github.com/marchaeologist/fabryka>. You will also find previous versions of the application archived there.

## TUTORIAL AND EXAMPLE DATASET

A tutorial is included in the application (*i.e.* 'Tutorial" panel) and can also be found in a dedicated article: <https://zenodo.org/records/15236695>.
'fabryka' includes an example file you can download from the application ('Upload data' subpanel).

## NOTICE

This project is licensed under the [MIT](https://choosealicense.com/licenses/mit/) License — Copyright (c) 2025 fabryka.

## CITATION

THOMAS, M. (2025). Fabryka: A web application for easily analysing & exploring the fabric of archaeological assemblages. In PCI archaeology (Version 1). Zenodo. https://doi.org/10.5281/zenodo.15236695